---
title: "SDC Data Visualization Guidelines"
subtitle: "Working document"
author: "Cedric Vidonne"
date: today
format:
  html:
    theme: [default, custom.scss]
  typst: 
    papersize: a4
---

```{r}
#| label: setup
#| include: false
source("R/palettes.R")
source("R/helpers.R")
```

## Colours

The palettes below are for **encoding data only** — bars, lines, pie slices, map areas. Colours for chart *elements* (axes, gridlines, labels) are handled
separately, later.

Every palette is shown at **7 colours** so the full set is visible. In practice a categorical palette stays readable up to about **5–6 categories**; beyond that, colours become hard to tell apart, including for colour-blind readers.

### Current palette

The categorical palette from the EDA strategy — our starting point and baseline for comparison.

```{r}
#| label: current-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "current")
```

### OKLCH

Built on a perceptual lightness staircase in OKLCH: same teal and pinkred anchors, with the light-pink and pure red swapped for a blue and a violet. 

**Pro:** even lightness, more colour-blind-robust, works in black & white. 

**Con:** slightly more muted — loses some of the current palette's .

```{r}
#| label: oklch-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "oklch")
```

### Sequentials

Sequential palettes based on the OKLCH approach for categorical.

```{r}
#| label: seq
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
library(dplyr)
library(purrr)
library(tibble)
library(chroma)

names <- c(
  "teal",
  "magenta",
  "orange",
  "yellow",
  "green",
  "blue",
  "purple"
)

oklch_seq <- palettes |>
  filter(palette == "oklch") |>
  mutate(name = names) |>
  mutate(
    colors = map(
      hex,
      ~ {
        interp_colors(
          9,
          colors = c("#121212", .x, "#f5f5f5"),
          correct.lightness = TRUE,
          model = "hcl"
        )[-c(1, 9)] # drop dark and light endpoints
      }
    )
  ) |>
  select(position, colors, name) |>
  tidyr::unnest_longer(colors) |>
  group_by(position) |>
  mutate(
    palette = paste0("seq_", name),
    shade = row_number(),
    hex = colors
  ) |>
  ungroup() |>
  select(palette, position = shade, hex)
```

#### Teal

```{r}
#| label: seq-teal
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_teal")
```

#### Magenta

```{r}
#| label: seq-magenta
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_magenta")
```

#### Orange

```{r}
#| label: seq-orange
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_orange")
```

#### Yellow

```{r}
#| label: seq-yellow
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_yellow")
```

#### Green

```{r}
#| label: seq-green
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_green")
```

#### Blue

```{r}
#| label: seq-blue
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_blue")
```

#### Purple

```{r}
#| label: seq-purple
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center

show_palette(oklch_seq, "seq_purple")
```



## Typography

## Layout

## Icons
