---
title: "SDC Data Visualization Guidelines"
subtitle: "Working document"
author: "Cedric Vidonne"
date: today
format:
  html:
    theme: [default, custom.scss]
    embed-resources: true
---

```{r}
#| label: setup
#| include: false
source("R/palettes.R")
source("R/helpers.R")
```

## Colours

The palettes below are for **encoding data only** — bars, lines, pie slices, map areas. Colours for chart *elements* (axes, gridlines, labels) are handled
separately, later.

Every palette is shown at **7 colours** so the full set is visible. In practice a categorical palette stays readable up to about **5–6 categories**; beyond that, colours become hard to tell apart, including for colour-blind readers.

### Current palette

The categorical palette from the EDA strategy — our starting point and baseline for comparison.

```{r}
#| label: current-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "current")
```

### OKLCH

Built on a perceptual lightness staircase in OKLCH: same teal and pinkred anchors, with the light-pink and pure red swapped for a blue and a violet. 

**Pro:** even lightness, more colour-blind-robust, works in black & white. 

**Con:** slightly more muted — loses some of the current palette's .

```{r}
#| label: oklch-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "oklch")
```

### Bright one

The same hues pushed higher in chroma for a livelier, near-neon feel. 

**Pro:** punchy and high-energy, strong on screen. 

**Con:** reads less institutional and can strain the eye in large filled areas.

```{r}
#| label: bright-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "bright")
```

### GPT one

A muted, earthy take: desaturated naturals (brown, gold, olive) alongside the teal and pinkred anchors. 

**Pro:** soft, cohesive, easy on the eye for large filled areas. 

**Con:** lower chroma makes some hues harder to tell apart at small sizes, and it feels less lively.

```{r}
#| label: gpt-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "gpt")
```

### Claude one

Similar earthy direction, but engineered so every colour clears a minimum contrast on white and the lightness steps stay ordered. 

**Pro:** accessible and well-balanced, dependable across chart types. 

**Con:** the darker, contrast-driven hues compress the range and can look a little sober. 

```{r}
#| label: claude-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "claude")
```

### Warm one

Leans warm: two or three neighbouring hot hues (brick, pinkred, orange, gold) before spreading to cool accents.

**Pro:** cohesive, warm identity. 

**Con:** the warm cluster risks similar lightness, reducing how many colours truly stand apart.

```{r}
#| label: warm-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "warm")
```

### Cool one

The mirror of warm — anchored in blues, greens and teal, with warm colours only as accents. 

**Pro:** calm and institutional. 

**Con:** the cool cluster can look samey and less energetic.

```{r}
#| label: cool-palette
#| echo: false
#| fig-width: 8
#| fig-height: 1.7
#| fig-align: center
show_palette(palettes, "cool")
```



## Typography

## Layout

## Icons
